### 2024-10-29 manuscript/Appendix_S4_v01.Rmd
Updated MC sims (across d, h2). Started walk-through on sim+hmsc tests across h2 levels.

---
title: "Appendix S4. The relative importance of evolution in populations and communities"
author:
  - Jelena H. Pantel^[Laboratoire Chrono-environnement,UMR 6249 CNRS-UFC, 16 Route de Gray, 25030 Besançon cedex, France, jelena.pantel@univ-fcomte.fr]
  - Ruben J. Hermann^[University of Duisburg-Essen, Universitätsstraße 5, 45141 Essen, Germany, ruben.hermann@uni-due.de]
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    toc: no
    fig_caption: yes
    latex_engine: xelatex
    extra_dependencies: ["float"]
  bookdown::html_document2: default
  word_document:
    reference_docx: docx_template.docx
header-includes:
- \usepackage{fancyhdr}
- \pagestyle{fancy}
- \fancyhead[L]{Author Pantel and Hermann}
- \fancyhead[R]{Detection of ecoevo dynamics}
- \usepackage{lineno}
- \linenumbers
link-citations: yes
linkcolor: blue
csl: the-american-naturalist.csl
bibliography: refs.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r setup,echo=FALSE,message=FALSE}
library(ecoevor)
library(ggplot2)
library(reshape2)
library(Hmsc)
library(bayesplot)
```
```{r,echo=FALSE,message=FALSE,warning=FALSE}
# Wrap long lines
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), tidy=TRUE,out.width="50%")
```

# Logistic growth and trait evolution

We can create simulations of evolutionary rescue in a population, where adaptive evolution rescues a population from extinction, at a speed that depends on the heritability level in the system. Population dynamics are modelled as:

$$ N_{i,t+1} =\frac{\hat{W}e^ \frac{-[(\frac{w+(1-h^2)P} {P+w})(E-x_{i,t})]^2}{2(P+w)}N_{i,t}} {1 + \alpha_{ii}N_{i,t} + \alpha_{ij} N_{j,t}} $$

where \( \hat{W} \) is calculated as \( \hat{W}=W_{max}\sqrt(\frac{w}{P+w}) \), \(W_{max}\) is the species' maximum per-capita growth rate, *w* is the width of the Gaussian fitness function (which determines the strength of selection, as increasing values indicate a weaker reduction in fitness with distance from optimum trait value), *P* is the width of the distribution of the phenotype *x*, and $h^2$ is the heritability of the trait *x*. For the simulation we use \(W_{max}\) = 2, *P* = 1, and *w* = 2.

The change in the average trait value each time step is given by:

$$ d_{i,t+1} = k d_{i,t} $$
where \( k = \frac{w + (1-h^2) P}{w + P} \) and \( d_{i,t} = E_t - x_{i,t} \).

## Population dynamics simulation

```{r}
### 1 species, h2=0 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- N
gdat$time <- 1:t
gdat$h2 <- h2
gdat$x1 <- x$x1
gdat$E <- E
```
```{r}
### 1 species, h2=0.01 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.01
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.02 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.02
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.03 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.03
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.04 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.04
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.05 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.05
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.075 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.075
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.1 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.1
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.4 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.4
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=0.9 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 0.9
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```
```{r}
### 1 species, h2=1 ###
# Initial conditions
N0 <- 10
alpha.11 <- 0.01
alpha <- matrix(alpha.11,byrow=FALSE)
E.0 <- 0.8
P <- 1
w <- 0.5
# Draw initial trait value using degree of initial maladaptation
B0 <- 1
d0 = sqrt(B0*(w+P))
x.0 = E.0-d0
Wmax <- 2
h2 <- 1
k <- (w + (1 - h2) * P)/(P + w)
# Simulation of model for t time steps
# Simulation of model for t time steps
t <- 60
N <- array(NA,dim=c(t,length(N0)))
N <- as.data.frame(N)
colnames(N) <- paste0("N",1:length(N0))
N[1,] <- N0
E <- rep(NA, t)
E[1] <- E.0
x <- array(NA,dim=c(t,length(N0)))
x <- as.data.frame(x)
colnames(x) <- paste0("x",1:length(N0))
x[1,] <- x.0
r <- array(NA,dim=c(t,length(N0)))
r <- as.data.frame(r)
colnames(r) <- paste0("r",1:length(N0))
What <- Wmax*sqrt(w/(P+w))
r[1,] <- What*exp((-(((w+(1-h2)*P)/(P+w))*(E[1]-x[1,]))^2)/(2*(P+w)))
for (i in 2:t) {
  res <- disc_LV_evol(N0=N[i-1,],alpha=alpha,E=E[i-1],x=x[i-1,],P=P,w=w,Wmax=Wmax,h2=h2)
  N[i,] <- res$Nt1
  r[i,] <- res$r
  # trait change
  d <- E[i-1] - x[i-1,]
  d1 <- k*d
  x[i,] <- E[i-1] - d1
  # environmental change
  E[i] <- E[i-1] + abs(rnorm(1, 0, 0))
}
# Arrange to save across scenarios
gdat <- rbind(gdat,data.frame(N,time=1:t,h2=rep(h2,t),x1=x$x1,E=E))
```

```{r}
### Plot of time series ###
gdat$h2 <- as.factor(gdat$h2)
ggplot2::ggplot(gdat, aes(time, N1, col=h2)) + geom_point()
### ###
```

## HMSC model fit

The goal of fitting the data to a statistical model is to estimate the relative important of trait evolution for the population dynamics.

```{r}
### Model 1, h2=00 ###
dat <- cbind(log(gdat$N1[gdat$h2==0]),gdat$x1[gdat$h2==0])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.1.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.1.sample <- sampleMcmc(m.1.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m1.post.hmsc <- convertToCodaObject(m.1.sample)
```

```{r warning=FALSE}
VP1 <- computeVariancePartitioning(m.1.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.1.sample,VP1,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.1.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

```{r}
### Model 2, h2=0.01 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.01]),gdat$x1[gdat$h2==0.01])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.2.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.2.sample <- sampleMcmc(m.2.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m2.post.hmsc <- convertToCodaObject(m.2.sample)
```

```{r warning=FALSE}
VP2 <- computeVariancePartitioning(m.2.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.2.sample,VP2,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.2.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

```{r}
### Model 3, h2=0.02 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.02]),gdat$x1[gdat$h2==0.02])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.3.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.3.sample <- sampleMcmc(m.3.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m3.post.hmsc <- convertToCodaObject(m.3.sample)
```

```{r warning=FALSE}
VP3 <- computeVariancePartitioning(m.3.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.3.sample,VP3,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.3.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

```{r}
### Model 4, h2=0.03 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.03]),gdat$x1[gdat$h2==0.03])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.4.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.4.sample <- sampleMcmc(m.4.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m4.post.hmsc <- convertToCodaObject(m.4.sample)
```

```{r warning=FALSE}
VP4 <- computeVariancePartitioning(m.4.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.4.sample,VP4,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.4.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```
```{r}
### Model 5, h2=0.04 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.04]),gdat$x1[gdat$h2==0.04])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.5.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.5.sample <- sampleMcmc(m.5.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m5.post.hmsc <- convertToCodaObject(m.5.sample)
```

```{r warning=FALSE}
VP5 <- computeVariancePartitioning(m.5.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.5.sample,VP5,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.5.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```
```{r}
### Model 6, h2=0.05 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.05]),gdat$x1[gdat$h2==0.05])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.6.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.6.sample <- sampleMcmc(m.6.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m6.post.hmsc <- convertToCodaObject(m.6.sample)
```

```{r warning=FALSE}
VP6 <- computeVariancePartitioning(m.6.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.6.sample,VP6,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.6.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

```{r}
### Model 7, h2=0.075 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.075]),gdat$x1[gdat$h2==0.075])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.7.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.7.sample <- sampleMcmc(m.7.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m7.post.hmsc <- convertToCodaObject(m.7.sample)
```

```{r warning=FALSE}
VP7 <- computeVariancePartitioning(m.7.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.7.sample,VP7,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.7.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```
```{r}
### Model 8, h2=0.1 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.1]),gdat$x1[gdat$h2==0.1])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.8.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.8.sample <- sampleMcmc(m.8.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m8.post.hmsc <- convertToCodaObject(m.8.sample)
```

```{r warning=FALSE}
VP8 <- computeVariancePartitioning(m.8.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.8.sample,VP8,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.8.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```
```{r}
### Model 9, h2=0.4 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.4]),gdat$x1[gdat$h2==0.4])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.9.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.9.sample <- sampleMcmc(m.9.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m9.post.hmsc <- convertToCodaObject(m.9.sample)
```

```{r warning=FALSE}
VP9 <- computeVariancePartitioning(m.9.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.9.sample,VP9,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.9.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```
```{r}
### Model 10, h2=0.9 ###
dat <- cbind(log(gdat$N1[gdat$h2==0.9]),gdat$x1[gdat$h2==0.9])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.10.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.10.sample <- sampleMcmc(m.10.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m10.post.hmsc <- convertToCodaObject(m.10.sample)
```

```{r warning=FALSE}
VP10 <- computeVariancePartitioning(m.10.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.10.sample,VP10,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.10.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```
```{r}
### Model 11, h2=1 ###
dat <- cbind(log(gdat$N1[gdat$h2==1]),gdat$x1[gdat$h2==1])
dat <- as.data.frame(dat)
colnames(dat) <- c("N1","x1")
dat$time <- 1:t
dat <- as.data.frame(dat)
df <- data.frame(dat[(2:t),-3])
colnames(df) <- c("Nt1", "xt1")
df$dx1 <- abs(dat$x1[2:t] - dat$x1[1:(t - 1)])
# prepare data in HMSC format
Y <- as.matrix(df$Nt1)
XData <- data.frame(cbind(E[1:(t-1)],E[1:(t-1)]^2,abs(dat$x1[2:t] - dat$x1[1:(t-1)])))
colnames(XData) <- c("E","Esq","dx1")

studyDesign = data.frame(sample = as.factor(1:(t-1)))
rL = HmscRandomLevel(units = studyDesign$sample)

m.11.hmsc = Hmsc(Y = Y, XData = XData, XFormula = ~E+Esq+dx1, studyDesign = studyDesign, ranLevels = list(sample = rL))
# Bayesian model parameters
nChains <- 2
thin <- 5
samples <- 2000
transient <- 1000*thin
verbose <- 500*thin
# sample MCMC
m.11.sample <- sampleMcmc(m.11.hmsc,thin=thin,sample=samples,transient=transient,nChains=nChains,verbose=verbose)
m11.post.hmsc <- convertToCodaObject(m.11.sample)
```

```{r warning=FALSE}
VP11 <- computeVariancePartitioning(m.11.sample,group=c(1,1,1,2),groupnames=c("Env","x1"))
plotVariancePartitioning(m.11.sample,VP11,cols=c("white","skyblue","darkgrey"),args.legend=list(cex=0.75,bg="transparent"))
```
```{r}
# prepare for plot
pred2 <- predict(m.11.sample)
hold1 <- list()
hold2 <- list()
for(i in 1:length(N0)){
  hold1[[i]] <- as.numeric(Y[,i])
  data <- lapply(pred2, function(x) x[, colnames(x)[i], drop = FALSE])
  hold2[[i]] <- matrix(unlist(data), nrow=length(pred2), byrow=TRUE)
}
y <- unlist(hold1) # vector of each species' observed N across time points
yrep <- as.matrix(data.frame(hold2))
### Mod2 plot ###
par(mgp=c(2,0.45,0), tcl=-0.4, mar=c(1.3,1.2,0,0))
color_scheme_set("brightblue")
#ppc_dens_overlay(y, yrep[1:50, ])
ppc_intervals(y,yrep,x=dat$time[1:(t-1)],prob = 0.95) +
  labs(x = "time",y = "ln(N)",) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
  theme(legend.position = "none")
```

```{r}
mc_dat_1 <- mcmc_intervals_data(m1.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_2 <- mcmc_intervals_data(m2.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_3 <- mcmc_intervals_data(m3.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_4 <- mcmc_intervals_data(m4.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_5 <- mcmc_intervals_data(m5.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_6 <- mcmc_intervals_data(m6.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_7 <- mcmc_intervals_data(m7.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_8 <- mcmc_intervals_data(m8.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_9 <- mcmc_intervals_data(m9.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_10 <- mcmc_intervals_data(m10.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")
mc_dat_11 <- mcmc_intervals_data(m11.post.hmsc$Beta, pars="B[dx1 (C4), sp1 (S1)]")

combined <- rbind(mc_dat_1,mc_dat_2,mc_dat_3,mc_dat_4,mc_dat_5,mc_dat_6,mc_dat_7,mc_dat_8,mc_dat_9,mc_dat_10,mc_dat_11)
combined$model <- rep(c("0","0.01","0.02","0.03","0.04","0.05","0.075","0.1","0.4","0.9","1"),times=c(dim(mc_dat_1)[1],dim(mc_dat_2)[1],dim(mc_dat_3)[1],dim(mc_dat_4)[1],dim(mc_dat_5)[1],dim(mc_dat_6)[1],dim(mc_dat_7)[1],dim(mc_dat_8)[1],dim(mc_dat_9)[1],dim(mc_dat_10)[1],dim(mc_dat_11)[1]))

# make the plot using ggplot 
theme_set(bayesplot::theme_default())
pos <- position_nudge(y = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1))
ggplot(combined, aes(x = m, y = parameter, color = model)) + 
  geom_linerange(aes(xmin = l, xmax = h), position = pos, size=2)+
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos)+
  geom_point(aes(color=model),position = pos)

ggplot(combined, aes(x = m, y = parameter, color = model)) + 
  geom_linerange(aes(xmin = l, xmax = h), position = pos, size=2)+
  geom_linerange(aes(xmin = ll, xmax = hh), position = pos)+
  geom_point(aes(color=model),position = pos)+
  coord_flip()+
  geom_vline(xintercept=0,linetype="dashed")
```

# References